Classical Simulation of Quantum Error Correction in a Fibonacci Anyon Code

### 2015-05-06 fib2.tex
Finished the discussion section.

Below the (bond) percolation threshold for (say) a 2D square lattice, we expect random sets of 
bonds to decompose into separate clusters~\footnote{Note that we use the word cluster to refer to both groupings of nearby anyons as formed by the decoding algorithm, and also to connected components of a graph under percolation, since both of these usages are standard in the literature. This distinction should be kept in mind, though context should remove any ambiguity regarding the meaning.} of average size $O(\log(n))$ and variance 
$O(1)$~\cite{Bazant2000}.
Each noise process in our model is associated with a (randomly distributed) edge, and so 
disconnected clusters correspond to sets of anyons that could not have interacted at any 
point in their history. As such, we can simulate the braiding processes of each cluster 
separately. In other words: the quantum state in the fusion space of all anyons factorizes into 
a tensor product over bond clusters, and we can compute each factor separately. Since each 
cluster has size only $O(\log(n))$, we can typically simulate these dynamics efficiently 
because the resulting fusion space has dimension $O(\mathrm{poly}(n))$. There are rare 
random processes that violate this reasoning, but processes such as these are suppressed 
exponentially in $L$, the lattice size~\cite{Grimmett1989}.

@@

Although we can guarantee that error-correction will succeed whenever noise does not percolate, it is possible that percolated events may still not result in an error. The connection between the percolation threshold and the error-correction threshold is not well understood in general~\cite{Hastings2014}, though it is clear that our threshold estimate will be a lower bound for the true threshold that may have been found if all events (including those that have percolated) were simulated. However, by simulating such percolated events on small lattices, our evidence suggests that for the Fibonacci anyon code that we study, the vast majority of such events will result in a logical error. As such we feel confident that neglecting such events does not have a significant effect on the threshold value obtained.

@@

We have demonstrated classical simulation of successful error correction in a universal anyon 
model. Though we have chosen several properties of our model and simulation in a 
convenient way for simplicity, we think it very unlikely that they will affect the qualitative 
results, similarly to the results of Ref.~\cite{Brell2013}. In particular, although we have 
modelled our logical qubit as encoded in the global topological degrees of freedom of our 
system, we could have encoded it in the fusion space of several preferred $\tau$ anyons. 
This situation would be appropriate to model error-correction routines in a topological 
quantum computation protocol. Additionally, we expect our results to be stable to changes 
in details of the noise model, decoding algorithm, and so on.

@@

Applying our methods to more realistic models such as concrete microscopic spin models or models with non-topological features is an open question.

\cggb{Comment on applicability of our simulation techniques to non-phenomenological 
models. Is it possible that the same ideas could be useful when there is more microscopic 
detail available, or are they restricted to exactly our phenomenological setup?}

\stf{Is there a possibility of using MPS like Rob Pfeifer to improve the simulations or 
include interactions?}~\cite{Pfeifer2010, Pfeifer2012, Singh2014}

\dude{Our simulation is effectively doing a contraction of a 2d peps-like network. So
yeah, maybe there are connections.}

\cggb{Do we have much more to say?}

>>>>

Below the (bond) percolation threshold for (say) a 2D square lattice, we expect random sets of 
bonds to decompose into separate clusters~\footnote{We use the word cluster to refer both to groupings of nearby anyons as formed by the decoding algorithm and to connected components of a graph under percolation, since both of these usages are standard in the literature. The intended meaning will remain clear from context.} of average size $O(\log(n))$ and variance 
$O(1)$~\cite{Bazant2000}.
Each noise process in our model is associated with a (randomly distributed) edge, and so 
disconnected clusters correspond to sets of anyons that could not have interacted at any 
point in their history. As such, we can simulate the braiding processes of each cluster 
separately. In other words: the quantum state in the fusion space of all anyons factorizes into 
a tensor product over bond clusters, and we can compute each factor separately. Since each 
cluster has size only $O(\log(n))$, we can typically simulate these dynamics efficiently 
because the resulting fusion space has dimension $O(\mathrm{poly}(n))$. There are rare 
random processes that violate this reasoning, but processes such as these are suppressed 
exponentially in $L$, the lattice size~\cite{Grimmett1989}.

@@

We can guarantee that error-correction will succeed whenever noise does not percolate, but it is possible that percolated events may still result in no error. The connection between the percolation threshold and the error-correction threshold is not well understood in general~\cite{Hastings2014}, though it is clear that our threshold estimate will be a lower bound for the true threshold that may be found if all events (including those that have percolated) were simulated by an optimal decoder. However, by simulating such percolated events on small lattices, our evidence suggests that for the Fibonacci anyon code that we study, the vast majority of such events will result in a logical error. As such it is likely that neglecting such events does not significantly effect our observed threshold value.

@@

We have demonstrated classical simulation of successful error correction in a universal anyon 
model. Though we have chosen several properties of our model and simulation in a 
convenient way for simplicity, we think it very unlikely that they will affect the qualitative 
results, similarly to the results of Ref.~\cite{Brell2013}. In particular, although we have 
modelled our logical qubit as encoded in the global topological degrees of freedom of our 
system, we could have encoded it in the fusion space of several preferred $\tau$ anyons. 
This situation would be appropriate to model error-correction routines for topological 
quantum computation. Additionally, we expect our results to be stable to changes 
in details of the noise model, decoding algorithm, and so on.

@@

There are several interesting avenues for further research. Applying these methods to more realistic models such as concrete microscopic spin models or models with non-topological features would give direct insight into practical error rates needed for topological quantum memories in non-Abelian systems. Another potential direction is to incorporate tensor network methods for simulating anyons~\cite{Pfeifer2010, Pfeifer2012, Singh2014} to improve the speed and scale of the simulations.
